---
title: 'Homework 2 - Research in Health Economics'
author: "Ammarah Ahmed"
date: ""
output: pdf_document 
---

```{r setup, include=FALSE}
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, stringr, readxl, data.table, gdata, readr,scales, kableExtra, lemon, imputeTS)
```

```{r, warning = FALSE, echo = FALSE, message = FALSE}
knit_print.data.frame <- lemon_print
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE, fig.width = 16/2, fig.height = 9/2, tidy.opts=list(width.cutoff=60), tidy=TRUE)
```

```{r, include=FALSE}
final.hcris.data<- readRDS("data/output/HCRIS_Data.rds")
```

#Question 1 

```{r, include=FALSE}
multiple_reports_table<- final.hcris.data %>% group_by(street, year) %>% 
  summarise(count = n()) %>%
  filter(count > 1) %>% 
  group_by(year) %>% 
  summarise(count = n())
```

```{r, include=FALSE}
multiple_reports<- sum(multiple_reports_table$count)
```

```{r}
paste0(multiple_reports , " hospitals filed more than one report per year. ")
```

```{r, include=FALSE}
fig.mult.rep<- multiple_reports_table %>% ggplot(aes(x = year, y = count)) +
  geom_line() + 
  labs(title = 'Number of Hospitals with Multiple Reports', x = 'Year', y = 'Number of hospitals') +
  theme_minimal()
```

```{r}
fig.mult.rep
```

#Question 2

```{r, include=FALSE}
medicare_id_table<- final.hcris.data %>% group_by(provider_number) %>% count() 
```
```{r, include=FALSE}
medicare_id<-nrow(medicare_id_table)
```

```{r}
paste0("There are ", medicare_id, " unique Medicare IDs")
```

#Question 3

```{r, include=FALSE}
tot_charges <- final.hcris.data %>% ggplot(aes(x = as.factor(year), y = tot_charges)) +
  geom_violin(trim = FALSE) +
  labs(title = 'Distribution of Total Charges', x = 'Year', y = 'Total Charges') +
  theme_minimal()
```

```{r}
tot_charges
```

#Question 4

```{r, include=FALSE}
final.hcris.data2<- final.hcris.data %>%
  mutate(discount_factor = 1-tot_discounts/tot_charges,
          price_num = (ip_charges + icu_charges + ancillary_charges)*discount_factor - tot_mcare_payment,
          price_denom = tot_discharges - mcare_discharges,
          est_price = price_num/price_denom)
```
```{r, include=FALSE}
est_prices<- final.hcris.data2 %>% 
  ggplot(aes(x = as.factor(year), y = est_price)) +
  geom_violin(trim = FALSE) + 
  labs(title = 'Distribution of Estimated Prices', x = 'Year', y = 'Estimated Prices') +
  theme_bw()
```

```{r}
est_prices
```

#Question 5

```{r, include=FALSE}
hcris_2012 <- final.hcris.data2 %>% filter(year == 2012)
```
```{r, include=FALSE}
hcris_2012$penalty <- ifelse(hcris_2012$hrrp_payment + hcris_2012$hvbp_payment < 0, 1,0)
```
```{r, include=FALSE}
pen_price<- hcris_2012 %>%
  filter(!is.na(penalty)) %>% 
  filter(penalty == 1) %>%
  group_by(penalty) %>% 
  summarise(price = mean(est_price, na.rm = TRUE))
```
```{r, include=FALSE}
non_pen_price <- hcris_2012 %>%
  filter(!is.na(penalty)) %>% 
  filter(penalty == 0) %>%
  group_by(penalty) %>% 
  summarise(price = mean(est_price, na.rm = TRUE))
```

```{r}
paste0("The average price amongst penalised hospitals is $", pen_price$price, " and the average price amongst non-penalised hospitals is $", non_pen_price$price)
```

#Question 6 

```{r, include=FALSE}
table_6 <- hcris_2012 %>% 
  group_by(street, penalty) %>% 
  summarise(avg_price = mean(est_price, na.rm = TRUE), quartile = quantile(beds, na.rm = TRUE)) %>%
  group_by(penalty) 

```
```{r}
options(knitr.kable.NA = 0)
knitr::kable(table_6, 
             col.names=c("Hospital","Penalty", "Average Price", "Quartile"),
             format.args = list(big.mark=","), caption = "Average Price for Treated and Control Groups")
```

#Question 7

```{r, include=FALSE}
hcris_2012_filt <- hcris_2012[!is.na(hcris_2012$est_price) & !is.na(hcris_2012$penalty) & !is.na(hcris_2012$quartile),]
```

```{r, include=FALSE}
m.inv.var <- Matching::Match(Y = hcris_2012_filt$est_price,
                             Tr = hcris_2012_filt$penalty,
                             X= hcris_2012_filt$quartile,
                             M = 1,
                             Weight = 1,
                             estimand = "ATE")
```
```{r}
summary(m.inv.var)
```

```{r, include=FALSE}
m.mahala <- Matching::Match(Y = hcris_2012_filt$est_price, 
                            Tr = hcris_2012_filt$penalty,
                            X = hcris_2012_filt$quartile,
                            M = 1,
                            Weight = 2,
                            estimand = "ATE")
```
```{r}
summary(m.mahala)
```

```{r, include=FALSE}
logit.model <- glm(penalty ~ beds + mcaid_discharges + ip_charges + mcare_discharges +
                     tot_mcare_payment, family=binomial, data=hcris_2012_filt)
ps <- fitted(logit.model)
```
```{r, include=FALSE}
m.inv.ps <- Matching::Match(Y=hcris_2012_filt$est_price,
                           Tr=hcris_2012_filt$penalty,
                           X=ps,
                           M=1,
                           estimand="ATE")
```

I get an error when using the inverse propensity weight method

```{r, include=FALSE}
reg1.dat <- hcris_2012_filt %>% filter(penalty == 1)
reg1 <- lm(est_price ~ beds, data=reg1.dat)
```
```{r, include=FALSE}
reg0.dat <- hcris_2012_filt %>% filter(penalty==0)
reg0 <- lm(est_price ~ beds, data=reg0.dat)
```
```{r, include=FALSE}
pred1 <- predict(reg1,new=hcris_2012_filt)
pred0 <- predict(reg0,new=hcris_2012_filt)
```
```{r}
mean(pred1-pred0)
```

I was unable to make a table as I got an error in one the above codes. 

#Question 8
I am unable to answer this question as some of my codes did not run 

#Question 9
I am unsure as I do not think my codes are correct

#Question 10
Working with this dataset was interesting. I learned more about applying the concepts we learned in class and was able to get more comfortable using R. However, it was confuding at times as I am still unclear about the codes to find ATE and struggled to write the codes to get to the required answers.





