---
title: 'Homework 2 - Research in Health Economics'
author: "Ammarah Ahmed"
date: ""
header-includes:
  - \usepackage{booktabs}
output: 
  bookdown::pdf_document2:
    fig_caption: yes
    toc: false
---

```{r setup, include=FALSE}
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, stringr, readxl, data.table, gdata, readr,scales, kableExtra, lemon, imputeTS)
```

```{r, warning = FALSE, echo = FALSE, message = FALSE}
knit_print.data.frame <- lemon_print
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE, fig.width = 16/2, fig.height = 9/2, tidy.opts=list(width.cutoff=60), tidy=TRUE)
```

```{r, include=FALSE}
load("Hw2_workpsace.Rdata")
```

```{r, include=FALSE}
final.hcris.data<- readRDS("data/output/HCRIS_Data.rds")
```

# Question 1 

```{r, include=FALSE}
multiple_reports_table<- final.hcris.data %>% group_by(street, year) %>% 
  summarise(count = n()) %>%
  filter(count > 1) %>% 
  group_by(year) %>% 
  summarise(count = n())
```

```{r, include=FALSE}
multiple_reports<- sum(multiple_reports_table$count)
```

`r multiple_reports` hospitals filed more than one report per year. 

```{r, include=FALSE}
fig.mult.rep<- multiple_reports_table %>% ggplot(aes(x = year, y = count)) +
  geom_line() + 
  labs(title = 'Number of Hospitals with Multiple Reports', x = 'Year', y = 'Number of hospitals') +
  theme_minimal()
```

```{r}
fig.mult.rep
```

# Question 2

```{r, include=FALSE}
medicare_id_table<- final.hcris.data %>% group_by(provider_number) %>% count() 
```
```{r, include=FALSE}
medicare_id<-nrow(medicare_id_table)
```

There are `r medicare_id` unique Medicare IDs

# Question 3

```{r, include=FALSE}
tot_charges <- final.hcris.data %>% ggplot(aes(x = as.factor(year), y = tot_charges, fill = "")) +
  geom_violin(trim = FALSE) + scale_y_continuous(trans = "log10") +
  labs(title = 'Distribution of Log of Total Charges', x = 'Year', y = 'Log of Total Charges') +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, size = 9, hjust = 1), plot.title = element_text(hjust = 0.5))
```

```{r}
tot_charges
```

# Question 4

```{r, include=FALSE}
final.hcris.data2<- final.hcris.data %>%
  mutate(discount_factor = 1-tot_discounts/tot_charges,
         price_num = (ip_charges + icu_charges + ancillary_charges)*discount_factor - tot_mcare_payment,
         price_denom = tot_discharges - mcare_discharges,
         est_price = price_num/price_denom)
```
```{r, include=FALSE}
final.hcris.data2 <- final.hcris.data2 %>%
  ungroup() %>% 
  filter(price_denom>100, !is.na(price_denom), 
         price_num>0, !is.na(price_num),
         est_price<100000, 
         beds>30) %>%
  mutate( hvbp_payment = ifelse(is.na(hvbp_payment),0,hvbp_payment),
          hrrp_payment = ifelse(is.na(hrrp_payment),0,abs(hrrp_payment)),
          penalty = (hvbp_payment-hrrp_payment<0))
```

```{r, include=FALSE}
est_prices<- final.hcris.data2 %>% 
  ggplot(aes(x = as.factor(year), y = est_price, fill = "")) +
  geom_violin(trim = FALSE) + 
  labs(title = 'Distribution of Estimated Prices', x = 'Year', y = 'Estimated Prices') +
  theme_bw()
```

```{r}
est_prices
```

# Question 5

```{r, include=FALSE}
hcris_2012 <- final.hcris.data2 %>% filter(year == 2012)
```
```{r, include=FALSE}
hcris_2012$penalty <- ifelse(hcris_2012$hvbp_payment - hcris_2012$hrrp_payment < 0, 1,0) 
```
```{r, include=FALSE}
pen_price<- hcris_2012 %>%
  filter(!is.na(penalty)) %>% 
  filter(penalty == 1) %>%
  group_by(penalty) %>% 
  summarise(price = mean(est_price, na.rm = TRUE))
```
```{r, include=FALSE}
non_pen_price <- hcris_2012 %>%
  filter(!is.na(penalty)) %>% 
  filter(penalty == 0) %>%
  group_by(penalty) %>% 
  summarise(price = mean(est_price, na.rm = TRUE))
```

The average price amongst penalised hospitals is `r round(pen_price$price)` and the average price amongst non-penalised hospitals is `r round(non_pen_price$price)`.


# Question 6 

```{r, include=FALSE}
hcris_2012$quartile <- ntile(hcris_2012$beds, 4)
```
```{r, include=FALSE}
hcris_2012$quartile_1 <- ifelse(hcris_2012$quartile == 1, 1,0)
hcris_2012$quartile_2 <- ifelse(hcris_2012$quartile == 2, 1,0)
hcris_2012$quartile_3 <- ifelse(hcris_2012$quartile == 3, 1,0)
hcris_2012$quartile_4 <- ifelse(hcris_2012$quartile == 4, 1,0)
```

```{r, include=FALSE}
table_6 <- hcris_2012 %>% filter(!is.na(penalty)) %>%
  group_by(quartile, penalty) %>%
  summarise(avg_price = mean(est_price, na.rm = TRUE))

```
```{r, include=FALSE}
table_6_2 <- pivot_wider(table_6, names_from = penalty, values_from = avg_price, 
                         names_prefix = "price")
```


```{r}
options(knitr.kable.NA = 0)
knitr::kable(table_6_2, 
             col.names=c("Quartile", "Average Price for Non-Penalised Hospitals", "Average Price for Penalised Hospitals"),
             format.args = list(big.mark=","), 
             caption = "Average Price for Penalised and Non-Penalised Hospitals",
             booktabs = TRUE) %>%
  kable_styling(latex_options = c("scale_down", "hold_position"))
```

# Question 7

```{r, include=FALSE}
m.inv.var <- Matching::Match(Y = hcris_2012$est_price,
                             Tr = hcris_2012$penalty,
                             X= hcris_2012 %>% select(quartile_1, quartile_2, quartile_3),
                             M = 1,
                             Weight = 1,
                             estimand = "ATE")
```
```{r, include=FALSE}
summary(m.inv.var)
```

```{r, include=FALSE}
m.mahala <- Matching::Match(Y = hcris_2012$est_price, 
                            Tr = hcris_2012$penalty,
                            X = hcris_2012 %>% select(quartile_1, quartile_2, quartile_3),
                            M = 1,
                            Weight = 2,
                            estimand = "ATE")
```
```{r, include=FALSE}
summary(m.mahala)
```

```{r, include=FALSE}
logit.model <- glm(penalty ~ quartile_1 + quartile_2 + quartile_3,
                   family = binomial, data = hcris_2012)

ps <- fitted(logit.model)
```
```{r, include=FALSE}
hcris_2012_2 <- hcris_2012 %>%
  mutate(ipw = case_when(
    penalty==1 ~ 1/ps,
    penalty==0 ~ 1/(1-ps),
    TRUE ~ NA_real_
  ))
```
```{r, include=FALSE}
mean.t1 <- hcris_2012_2 %>% filter(penalty==1) %>%
  select(est_price, ipw) %>% summarize(mean_p=weighted.mean(est_price,w=ipw))
```
```{r, include=FALSE}
mean.t0 <- hcris_2012_2 %>% filter(penalty==0) %>%
  select(est_price, ipw) %>% summarize(mean_p=weighted.mean(est_price,w=ipw))
```
```{r, include=FALSE}
inv.ps<- mean.t1$mean_p - mean.t0$mean_p
```

```{r, include=FALSE}
reg1.dat <- hcris_2012 %>% filter(penalty == 1)
reg1 <- lm(est_price ~ quartile_1 + quartile_2 + quartile_3, data=reg1.dat)
```
```{r, include=FALSE}
reg0.dat <- hcris_2012 %>% filter(penalty == 0)
reg0 <- lm(est_price ~ quartile_1 + quartile_2 + quartile_3, data=reg0.dat)
```
```{r, include=FALSE}
pred1 <- predict(reg1,new=hcris_2012)
pred0 <- predict(reg0,new=hcris_2012)
```
```{r, include=FALSE}
reg_res <- mean(pred1-pred0)
```

```{r, include=FALSE}
table_ate <- data.frame(Estimators = c("NN Matching, inverse variance", "NN Matching, mahalanobis",
                                       "Inverse pscore weighting", "Regression"),
                        ATE = c(m.inv.var$est, m.mahala$est, inv.ps, reg_res))
```

```{r}
options(knitr.kable.NA = 0)
knitr::kable(table_ate, 
             col.names=c("Estimate", "ATE"),
             format.args = list(big.mark=","), caption = "Average Treatment Effect")
```
 

# Question 8

The results from the different treatment effect estimators are exactly the same at 199.5281. This is likely because of the use of dummy variables to define the quartiles based on hospital bed sizes.    

# Question 9

I do not think the results here show a causal effect of penalty on price. This is because there could be other factors affecting hospital prices which have not been taken into consideration in this test. 

# Question 10

Working with this dataset was interesting. I learned more about applying the concepts we learned in class and was able to get more comfortable using R. However, it was confusing at times as I am still unclear about the codes to find ATE and struggled to write the codes to get to the required answers.

